% intro talk in german
%\documentclass[NET,a4paper,12pt,ngerman]{netforms}

% intro talk in english
\documentclass[NET,a4paper,12pt,english]{netforms}
\usepackage{hyperref}

\usepackage[utf8]{inputenc}
\usepackage{tumlang}
\usepackage{tumcontact}
\usepackage{scrlayer-scrpage}
\usepackage[textsize=scriptsize]{todonotes}
% Disable todonotes with
%\setuptodonotes{disable}
\setuptodonotes{inline}
\usepackage{pgfgantt}
\usepackage{float}
\usepackage{enumitem}
\newlist{researchquestions}{enumerate}{1}
\setlist[researchquestions,1]{label=\textbf{\textit{RQ \arabic*}}, resume}


\geometry{%
	top=20mm,
	bottom=20mm,
	left=25mm,
	right=25mm,
	headsep=1.5cm,
	includehead,
}
\input{studentdefs}

\pagestyle{scrheadings}
\clearscrheadfoot
\chead{\TUMheader{1cm}}

\renewcommand{\maketitle}{%
	\begin{center}
		\textbf{\introductoryheadline}%

		\Large%
		\textbf{\thetitle}%
	\end{center}

	\footnotesize%
	\hrule
	\vskip1ex
	\begin{tabular}{ll}
		\thenamelabel: & \thevorname{} \textbf{\thenachname}\\
		\theadvisorlabel: & \hspace*{-.5ex}\thebetreuer\\
		\thesupervisorlabel: & \chairhead\\
		\thebeginlabel: & \thebeginnmonat/\thebeginnjahr\\
		\theendlabel: & \theendmonat/\theendjahr\\
	\end{tabular}
	\vskip1ex
	\hrule
	\vskip4ex
}

\linespread{1.2}
\setlength{\parskip}{.5\baselineskip}

\begin{document}
\maketitle

\subsection*{Topic}

In this IDP project we will investigate Starlink internals. Starlink is a project from SpaceX~\cite{starlink} aiming to provide fast Internet connection
anywhere in the world. Our scope is to understand internal routing decisions, general behavior for handling incoming packets and to
comprehend how satellite handovers are performed.
Our plan plan to fuzz the dish internal gRPC API to see if we manage to get access to internal statistics, additional information will be extracted by dissecting the companion Android application.
Eventually we will develop a toolset to interact with the API, in order to abstract away some of the logic with the end goal of making it easier to perform managements. 
Additional tooling may be developed to understand how satellite handovers are made. 
We will also investigate the Starlink network behavior in order to try to make sense of sudden ping or bandwidth changes when doing measurements.

Starlink is an interesting project because it can provide Internet to remote areas and could potentially be a revolution in how we connect to the Internet, a huge part of the reasons why this project is getting so popular is that it is able to provide a satellite-based low latency connection. Being able to measure performance of this new technology could be very important and potentially we believe that our work can be important for further analysis.
\subsection*{Related Work}

There is not much literature about internals of the dish and about the API, there's an interesting hardware-level hacking talk given at DEFCON~\cite{defcon}, in which they managed to get a shell on the dish, however this hardware-level research won't be part of our research.

We found some research regarding measurements~\cite{10.1145/3517745.3561416, 10.1145/3517745.3561457}, in the latter study a consistent loss of packages correlated with satellite handovers was observed. We will try to provide some insights regarding the impact of satellite handovers in latency increases.

Additionaly there is a repository on GitHub from the Starlink community containing pre-generated Go gRPC bindings \cite{starlink-grpc-go}.

\subsection*{Research Questions}
\begin{researchquestions}
\item Are there any hidden API endpoints?
\item Can the dish data be used to deduct internal starlink network behavior?
\item Can we extract some data that helps us verify the data the API is offering?
\end{researchquestions}

\subsection*{Approach}
In the first part we will try to understand and document the API, then, we will work on a library to make access to the API easier.
We will develop this in Python, because of the great support of other libraries and because it seems to be the de-facto standard language
for data analysis and measurements. This will simplify data gathering for measurements performed by other researchers.

We will intercept API calls made by the Android application using a proxy like Burpsuite Community Edition, and will eventually dissect the APK file
using tools like Jadx or Apktool. If needed we can also use Frida to do some dynamic patching.

Finally, we combine the information from the API and dish with passive measurement data we extract to check the quality of the measurements offered by the API.
We will probably need to develop some tooling to extract the data and combine or compare the two data sources.


\subsection*{Planned Schedule}
  \resizebox{\textwidth}{!}{
  \begin{ganttchart}[
      x unit=0.7mm,
      y unit title=0.6cm,
      title height=1,
      title label font=\small,
      newline shortcut=true,
      bar label font=\scriptsize,
      milestone label font=\scriptsize,
      milestone inline label node/.append style={left=2mm},
      milestone/.append style={fill=orange, shape=rectangle},
      y unit chart=.4cm,
      time slot format=isodate,
      time slot unit=day,
      bar height=0.7,
    ]{2023-03-01}{2023-08-31}
    \gantttitlecalendar{month=name}\\
    \ganttbar{Explore and document dish API}{2023-03-01}{2023-03-31}\\
    \ganttbar{Dissect the Android application}{2023-04-01}{2023-04-31}\\
    \ganttbar{Develop  some tooling to interact with API }{2023-05-01}{2023-05-31}\\
    \ganttbar{Use the tool to understand internals}{2023-06-01}{2023-08-31}
  \end{ganttchart}
  \label{fig:work-plan}
  }

  \subsection*{Lecture}
  
  To complement my IDP, I chose \href{https://www.ce.cit.tum.de/en/lkn/lehre/vorlesungen/analysis-modeling-and-simulation-of-communication-networks}{"Analysis, Modeling and Simulation of Communication Networks"}. The IDP project work builds upon the foundations thought on the couse. The covered topics (e.\,g., packet level simulation and topology models) will help to understand routing decisions potentially taken by the dish. Moreover, the practical approach to the arguments will be useful for the implementation of solutions to test performance and behaviour of the dish while using Starlink.

\bibliographystyle{IEEEtran}
\scriptsize
\bibliography{IEEEabrv,lit}

\end{document}
